Gut microbes and their genes or something else

### 2023-02-17 manuscript/tex/sections/results.tex
clean up

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{multirow}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\usepackage[normalem]{ulem}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\hypertarget{title}{%
\subsection{Title}\label{title}}

\begin{itemize}
\item
  \begin{quote}
  Gut microbes and their genes are associated with cognition and
  neuroanatomy in children
  \end{quote}
\item
  \begin{quote}
  Gut microbes and their genes are associated with cognition performance
  and brain structure in children
  \end{quote}
\item
  \begin{quote}
  Extensive relationships between the gut microbiota and normal
  neurocognitive development
  \end{quote}
\item
  \begin{quote}
  Extensive relationships between the gut microbiota and typical
  neurocognitive development
  \end{quote}
\item
  \begin{quote}
  The gut microbial metagenome is connected to neurocognition in typical
  child development
  \end{quote}
\item
  \begin{quote}
  The microbial potential in regulating early childhood neurocognitive
  development
  \end{quote}
\end{itemize}

\hypertarget{authors}{%
\subsubsection{Authors}\label{authors}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  Bonham 0000-0003-3200-7533
  \end{quote}
\item
  \begin{quote}
  Bottino 0000-0003-1953-1576
  \end{quote}
\item
  \begin{quote}
  McCann 0000-0002-9753-7968
  \end{quote}
\item
  \begin{quote}
  Beauchemin \{ORCID??\}
  \end{quote}
\item
  \begin{quote}
  Elizabeth Weisse \textless elizabeth.weisse2@gmail.com\textgreater{}
  \end{quote}
\item
  \begin{quote}
  \ldots{}
  \end{quote}
\item
  \begin{quote}
  Jonathan O\textquotesingle Muircheartaigh
  \textless{}\href{mailto:jonathanom@kcl.ac.uk}{\uline{jonathanom@kcl.ac.uk}}\textgreater{}
  \end{quote}
\item
  \begin{quote}
  Huttenhower 0000-0002-1110-0096
  \end{quote}
\item
  \begin{quote}
  Muriel Bruchhage
  \textless{}\href{mailto:mmk.bruchhage@gmail.com}{\uline{mmk.bruchhage@gmail.com}}\textgreater{}
  0000-0001-9637-0951
  \end{quote}
\item
  \begin{quote}
  D'Sa 0000-0002-3476-8419
  \end{quote}
\item
  \begin{quote}
  Deoni 0000-0002-0633-6328
  \end{quote}
\item
  \begin{quote}
  Klepac-Ceraj 0000-0001-5387-5706
  \end{quote}
\end{enumerate}

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

The gastrointestinal tract, its resident microorganisms and the central
nervous system are connected by biochemical signaling, also known as
"microbiome-gut-brain-axis." Both the human brain and the gut microbiome
have critical developmental windows in the first three years of life,
raising the possibility that their development is co-occurring and
likely co-dependent. Emerging evidence implicates gut microorganisms and
microbiota composition in cognitive outcomes and neurodevelopmental
disorders (e.g., autism and anxiety), but the influence of gut microbial
metabolism on typical neurodevelopment has not been explored in detail.
We investigated the relationship of the microbiome with the neuroanatomy
and cognitive function of 361 healthy children, demonstrating that
differences in gut microbial taxa and gene functions are associated with
overall cognitive function as well as with differences in the size of
multiple brain regions. Using a combination of multivariate linear
models and machine learning (ML) models, we showed that many species,
including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
significantly associated with higher cognitive function, while some
species such as \emph{Ruminococcus gnavus} were more commonly found in
children with low cognitive scores regardless of age or maternal
education. Microbial genes for enzymes involved in the metabolism of
neuroactive compounds, particularly short-chain fatty acids such as
glutamate and propionate, were also associated with cognitive function.
In addition, ML models were able to use microbial taxa to predict the
volume of brain regions, with particular taxa often dominating the
feature importance metric for individual brain regions, such as \emph{B.
wexlerae} being the most dominant for the parahippocampal region or
GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
findings provide potential biomarkers of neurocognition and may lead to
the future development of targets for early detection and early
intervention.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The gut and the brain are intimately linked. Signals from the brain
reach the gut through the autonomic nervous system and the endocrine
system, and the gut can communicate with the brain through the vagus
nerve and through endocrine and immune (cytokine) signaling molecules
\href{https://www.zotero.org/google-docs/?i3SRPn}{(Cerdó et al., 2019;
Pronovost and Hsiao, 2019; Sharon et al., 2016; Tognini, 2017)}. In
addition, the products of microbial metabolism generated in the gut can
influence the brain, both indirectly by stimulating the enteric nervous
and immune systems, as well as directly through molecules that enter
circulation and cross the blood-brain barrier. Causal links between the
gut microbiome and neural development, particularly atypical
development, are increasingly being identified
\href{https://www.zotero.org/google-docs/?UsrFKB}{(Spichak et al.,
2021)}. Both human epidemiology and animal models point to the effects
of gut microbes on the development of autism spectrum disorder
\href{https://www.zotero.org/google-docs/?Y8l5A0}{(Laue et al., 2020;
Wan et al., 2021)} and particular microbial taxa have been associated
with depression
\href{https://www.zotero.org/google-docs/?mbaFE3}{(Mayneris-Perxachs et
al., 2022; Valles-Colomer et al., 2019)} and Alzheimer's disease
\href{https://www.zotero.org/google-docs/?0BEGUn}{(Fung et al., 2017;
Kim et al., 2021)}. But information about this ``microbiome-gut-brain
axis'' in normal neurocognitive development remains lacking,
particularly early in life.

The first years of life are critical developmental windows for both the
microbiome and the brain
\href{https://www.zotero.org/google-docs/?XTqhlm}{(Laue et al., 2022)}.
Fetal development \emph{in utero} is believed to be mostly sterile, but
is rapidly seeded at birth through contact with the birth canal (if
birthed vaginally), caregivers, food sources (breastmilk or formula),
and other environmental sources such as antibiotics
\href{https://www.zotero.org/google-docs/?X6N0f4}{(Bäckhed et al., 2015;
Bokulich et al., 2016; Dominguez-Bello et al., 2010; Louwies et al.,
2020)}. The early microbiome is characterized by low microbial
diversity, rapid succession and evolution, and domination by
Actinobacteria, particularly the genus \emph{Bifidobacterium},
Bacteroidetes, especially \emph{Bacteroides}, and Proteobacteria
\href{https://www.zotero.org/google-docs/?VG43uh}{(Koenig et al.,
2011)}. Many of these microbes have specialized metabolic capabilities
for digesting human breast milk, such as \emph{Bifidobacterium infantis}
and \emph{Bacteroides fragilis}
\href{https://www.zotero.org/google-docs/?1dKzKj}{(Sela et al., 2008;
Tso et al., 2021)}. Upon the introduction of solid foods, the gut
microbiome undergoes another categorical transformation, with most taxa
of the infant microbiome being replaced by taxa more reminiscent of
adult microbiomes
\href{https://www.zotero.org/google-docs/?V3lY3p}{(Bäckhed et al.,
2015)}. Many prior studies have focused on either infant microbiomes or
adult microbiomes, since performing statistical analyses across this
transition poses particular challenges. Nevertheless, since this
transition coincides with critical neural developmental windows and
neural synaptogenesis, investigation across this solid-food boundary is
incredibly important
\href{https://www.zotero.org/google-docs/?d29bv8}{(Tau and Peterson,
2010)}.

A child's brain undergoes remarkable anatomical, microstructural,
organizational, and functional changes in the first years of life. By
age 5, a child's brain has reached \textgreater85\% of its adult size,
has achieved near-adult levels of myelination, and the pattern of axonal
connections has been established
\href{https://www.zotero.org/google-docs/?1jUZOO}{(Silbereis et al.,
2016)}. Much of this development occurs in discrete windows or
``sensitive-'' or ``critical periods'' (CPs)
\href{https://www.zotero.org/google-docs/?w4Pcze}{(Knudsen, 2004)} when
neural plasticity is particularly high, and particular modes of learning
and skill development are preferred. The timing of these sensitive
periods is driven in part by genetics, but can also be affected by the
environment, including the gut microbiome
\href{https://www.zotero.org/google-docs/?KTm9lQ}{(Cowan et al., 2020)}.
In fact, emerging evidence suggests that the timing and duration of CPs
may be driven in part by cues from the developing gut microbiome
\href{https://www.zotero.org/google-docs/?mAD1ne}{(Callaghan, 2020)}. As
such, understanding the normal spectrum of healthy microbiome
development and how it relates to normal neurocognitive development may
provide opportunities for identifying atypical development earlier and
offer opportunities for intervention.

To begin to address this need, we investigated the gut microbiome and
neurocognitive development of children from infancy through 10 years of
age in an accelerated longitudinal study design. Gut microbial
communities were assessed using shotgun metagenomic sequencing, enabling
profiling at both the taxonomic and gene-functional level, and
neurocognitive development was measured using expert-assessment of
cognitive function and neuroimaging using magnetic resonance imaging
(MRI). Using a combination of classical statistical analysis and machine
learning, we showed that the development of the gut microbiome, the
brain, and children's cognitive abilities are intimately linked, with
both microbial taxa and gene functions able to predict cognitive
performance and brain structure.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{cohort-overview-and-summary-data}{%
\subsubsection{Cohort overview and summary
data}\label{cohort-overview-and-summary-data}}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2620}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1845}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Group}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Subgroup}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{All}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Under 6 mo}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Over 18 mo}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
N subjects
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
361
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
74
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
257
\end{minipage} \\
\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Samples
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
262 (72.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
74 (100.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
203 (79.0\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
72 (19.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
46 (17.9\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
\textgreater2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
27 (7.5\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
8 (3.1\%)
\end{minipage} \\
\multirow{3}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Age (months)
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
Min
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
18.0
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Max
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
119.3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6.0
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
119.3
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Median
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
25.8
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
3.7
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
45.9
\end{minipage} \\
\multirow{2}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Sex
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
F
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
163 (45.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (50.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
114 (44.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
M
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
198 (54.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (50.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
143 (55.6\%)
\end{minipage} \\
\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Race
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
White
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
252 (69.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
42 (56.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
191 (74.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Black
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
40 (11.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16 (21.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
21 (8.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Asian
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Mixed
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
7 (1.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (1.4\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6 (2.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Other
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
48 (13.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12 (16.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
29 (11.3\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Unknown/Declined
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2 (0.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (1.4\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
2 (0.8\%)
\end{minipage} \\
\multirow{6}{*}{\begin{minipage}[b]{\linewidth}\raggedright
Maternal Ed.
\end{minipage}} & \begin{minipage}[b]{\linewidth}\raggedright
Junior high school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0 (0.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
1 (0.4\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Some high school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
6 (1.7\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
3 (4.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
4 (1.6\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
High school grad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
37 (10.2\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
13 (17.6\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16 (6.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Some college
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
95 (26.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
25 (33.8\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
57 (22.2\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
College grad
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
90 (24.9\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
17 (23.0\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
69 (26.8\%)
\end{minipage} \\
& \begin{minipage}[b]{\linewidth}\raggedright
Grad/professional school
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
123 (34.1\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
15 (20.3\%)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
104 (40.5\%)
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

@@

\hypertarget{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}{%
\subsubsection{Microbial species and neuroactive genes are associated
with cognitive
performance}\label{microbial-species-and-neuroactive-genes-are-associated-with-cognitive-performance}}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2300}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1259}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1525}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1162}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2785} + 2\tabcolsep}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{All}
\end{minipage}} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2591} + 2\tabcolsep}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{under 6 months}
\end{minipage}} &
\multicolumn{2}{>{\raggedright\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.2324} + 2\tabcolsep}@{}}{%
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{over 18 months}
\end{minipage}} \\
\begin{minipage}[b]{\linewidth}\raggedright
geneset
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{E.S.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{q value}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1223
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5549
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1944
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1533}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1269
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6023
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. II
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2201
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0660}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3424
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0198}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0946
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9199
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Acetate syn. III
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2104
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0859}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3098
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0518}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0857
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9199
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Glutamate deg. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.4037
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1802}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2840
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.2489
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6023
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Glutamate syn. II
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0953
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.6022
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1342
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9802
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.2421
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0466}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Menaquinone syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0853
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5635
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0833
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.1699
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1825}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Propionate deg. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.4612
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.1302}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.5421
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0198}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6738
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0414}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Propionate syn. I
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3344
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5549
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.3201
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.5713
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.6760
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0227}
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
Tryptophan syn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.0573
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.9742
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.0725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.8159
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.1186
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{0.0414}
\end{minipage} \\
\midrule()
\endhead
\bottomrule()
\end{longtable}

\hypertarget{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}{%
\subsubsection{Gut microbial taxonomic and functional profiles predict
cognitive
function}\label{gut-microbial-taxonomic-and-functional-profiles-predict-cognitive-function}}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1456}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.0960}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3584}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Subject Ages (months)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Microbial feature
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Demo.
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Test set correlation (± C.I.)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Test set Root-mean-square error (± C.I.)
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.14 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
13.01 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.10 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.66 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
taxa
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.13 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.70 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.01 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.56 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
0 to 6
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
genes
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-0.01 ± 0.01
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
12.56 ± 0.05
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
0.506 ± 0.003
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
16.27 ± 0.04
\end{minipage} \\
\begin{minipage}[b]{\linewidth}\raggedright
18 to 120
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright

>>>>

\subsubsection{Cohort overview and summary data}

@@

@@

\subsubsection{Microbial species and neuroactive genes are associated with cognitive performance}

@@

\subsubsection{Gut microbial taxonomic and functional profiles predict cognitive function}

@@
